DRAFT Addition re Other Packages

### 2024-09-24 scratch/JSS-submission/Revision-1/Fox-Monette-cv-package.Rmd
add section on other software to JSS paper

---
documentclass: jss
author:
  - name: John Fox
    orcid: 0000-0002-1196-8012
    affiliation: 'McMaster University'
    address: |
      | Hamilton, Ontario, Canada
    email: \email{jfox@mcmaster.ca}
    url: https://www.john-fox.ca/
  - name: Georges Monette
    orcid: 0000-0003-0076-5532
    affiliation: 'York University'
    address: |
      | Toronto, Ontario, Canada
    email: \email
    url: http://blackwell.math.yorku.ca/gmonette
title:
  formatted: "\\pkg{cv}: An \\proglang{R} Package for Cross-Validating Regression Models"
  # If you use tex in the formatted title, also supply version without
  plain:     "cv: An R Package for Cross-Validating Regression Models"
  # For running headers, if needed
  short:     "\\pkg{cv}: Cross-Validating Regression Models"
abstract: >
  We describe the \pkg{cv} package, which implements cross-validation for standard \proglang{R} regression models through a uniform and simple to use generic function. Methods for the \code{cv()} function are provided for many commonly employed classes of statistical models, including mixed-effects models, and it is straightforward to extend the \pkg{cv} package by writing methods for additional classes of models. The \code{cv()} function can also cross-validate complex model-selection procedures, such as those that include variable transformations, predictor selection, and selection among competing models. The \code{cv()} function generally supports parallel computations, and the supplied methods for linear and generalized linear models take advantage of computational efficiencies.
keywords:
  # at least one keyword must be supplied
  formatted: [cross-validation, regression analysis, model selection, "\\proglang{R}"]
  plain:     [cross-validation, regression analysis, model selection, R]
preamble: >
  \usepackage{amsmath}
output: 
  rticles::jss_article:
    extra_dependencies: "subfig"
bibliography: ["cv.bib"]
---

@@

## Other statistical software: SAS, Stata, and Python

>>>>

---
documentclass: jss
author:
  - name: John Fox
    orcid: 0000-0002-1196-8012
    affiliation: 'McMaster University'
    address: |
      | Hamilton, Ontario, Canada
    email: \email{jfox@mcmaster.ca}
    url: https://www.john-fox.ca/
  - name: Georges Monette
    orcid: 0000-0003-0076-5532
    affiliation: 'York University'
    address: |
      | Toronto, Ontario, Canada
    email: \email
    url: http://blackwell.math.yorku.ca/gmonette
title:
  formatted: "\\pkg{cv}: An \\proglang{R} Package for Cross-Validating Regression Models"
  # If you use tex in the formatted title, also supply version without
  plain:     "cv: An R Package for Cross-Validating Regression Models"
  # For running headers, if needed
  short:     "\\pkg{cv}: Cross-Validating Regression Models"
abstract: >
  We describe the \pkg{cv} package, which implements cross-validation for standard \proglang{R} regression models through a uniform and simple to use generic function. Methods for the \code{cv()} function are provided for many commonly employed classes of statistical models, including mixed-effects models, and it is straightforward to extend the \pkg{cv} package by writing methods for additional classes of models. The \code{cv()} function can also cross-validate complex model-selection procedures, such as those that include variable transformations, predictor selection, and selection among competing models. The \code{cv()} function generally supports parallel computations, and the supplied methods for linear and generalized linear models take advantage of computational efficiencies.
keywords:
  # at least one keyword must be supplied
  formatted: [cross-validation, regression analysis, model selection, "\\proglang{R}"]
  plain:     [cross-validation, regression analysis, model selection, R]
preamble: >
  \usepackage{amsmath}
  \usepackage{amssymb}
output: 
  rticles::jss_article:
    extra_dependencies: "subfig"
bibliography: ["cv.bib"]
---

@@

## Comparing cv to other software for cross-validation: SAS, Stata, and Python

Although we do not have recent experience with statistical software for cross-validation other than in \proglang{R},[^other-software] we have examined provisions for CV in \proglang{SAS}, \proglang{Stata}, and \proglang{Python}, primarily by locating and reading documentation.

[^other-software]: It is our impression that these three statistical systems, in addition to \proglang{R}, are most likely to be used for applications of cross-validation. We were once frequent users of \proglang{SAS}, though not recently, and have only cursory experience with \proglang{Stata} and \proglang{Python}, so we are not ideally situated to evaluate CV software for any of the three.

### SAS

Cross-validation in \proglang{SAS} is associated with particular procedures ("\code{PROC}s"), and is distributed across several products, including, for example, \code{PROC LOGISTIC} and \code{PROC PLS} in \proglang{SAS/STAT} [@SAS-Stat:2020], in \proglang{SAS Enterprise Miner} [@SAS-Enterprise-Miner:2018], and in \proglang{SAS Visual Data Mining and Machine Learning} [@SAS-Visual-Data-Mining:2020]. These appear to be straightforward implementations of $k$-fold CV.

In addition, \proglang{SAS} has programming capabilities, including via its macro language, and we're aware of several independent implementations of CV in \proglang{SAS} macros. These tend to be limited in scope. For example @Slaets-et-al:2014 employ and provide access to a \proglang{SAS} macro for $k$-fold CV of mixed-effects models, which repeatedly calls \code{PROC MIXED} in \proglang{SAS/STAT}, and which performs cluster-based resampling (one of the options that we provide for mixed-effects models in the \pkg{cv} package).

### Stata

As far as we can tell, \proglang{Stata} has no standard implementation of cross-validation. Like \proglang{R}, however, \proglang{Stata} is extensible through a programming language that allows users to define new commands in so-called "\code{ado}" files.

Several such contributed \proglang{Stata} CV commands, varying in generality, are shared in an archive maintained at the Boston College Department of Economics. For example, the \code{CV\_REGRESS} command [@Rios-Avila:2018] computes leave-one-out CV for least-squares regression using the hatvalues, and thus is comparable to \code{cv()} for linear models with \code{method = "hatvalues"}. In contrast, the \code{CROSSVALIDATE} command [@Schonlau:2020] performs $k$-fold CV for a wide range of regression models fit in \proglang{Stata}, and is roughly comparable to the default method of \code{cv()}, although \code{CROSSVALIDATE} returns cross-validated fitted values, from which cost criteria can be computed, rather than (as \code{cv()}) cost-criteria directly.

### Python

The programming language \proglang{Python} is often the software of choice for "machine learning," in which cross-validation can play a prominent role. Machine learning and statistical modeling more generally are supported through extensive \proglang{Python} libraries, such as \pkg{scikit-learn} [@Pedregosa-et-al:2011; @ScikitLearnDevelopers:2024] and \pkg{TensorFlow} [@TensorFlow:2015, which isn't limited to use within \proglang{Python}].

Functions for cross-validation are provided in \pkg{scikit-learn}, which supports $k$-fold, repeated $k$-fold, and leave-one-out CV. These functions can be applied to data sets consisting of independent observations, hierarchical groups, or time-series data.  A sample program in the \pkg{scikit-learn} documentation [@ScikitLearnDevelopers:2024] demonstrates how to implement meta cross-validation (termed "nested cross-validation" in the documentation).

The approach taken by \pkg{scikit-Learn} for independent observations is similar to that in the default \code{cv()} function in the \pkg{cv} package. The approach for hierarchical data is similar, but not identical, to that in \pkg{cv} (as described in Section \ref{cross-validating-mixed-effects-models}). Our \code{cv()} package doesn't currently provide for cross-validating time-series models.
